---
title: "ehlan_siddiqi_midterm"
author: "Ehlan Siddiqi"
date: "7/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
```

[GitHub - Click here!](https://github.com/ehlan/compscix-415-2-assignments)

## Midterm Examination - Ehlan Siddiqi
### Table of Content (click to jump to each section)
<a href="#sec1">1. RStudio and R Markdown</a><br>
<a href="#sec2">2. The tidyverse packages</a><br>
<a href="#sec3">3. R Basics</a><br>
<a href="#sec4">4. Data import/export</a><br>
<a href="#sec5">5. Visualization</a><br>
<a href="#sec6">6. Data munging and wrangling</a><br>
<a href="#sec7">7. EDA</a><br>
<a href="#sec8">8. GIT and Github</a><br>


##<a name="sec1">1.</a> RStudio and R Markdown
 - Provided
 
##<a name="sec2">2.</a> The tidyverse packages

**1. Can you name which package is associated with each task below?**

&nbsp;&nbsp;&nbsp;**Plotting -** ggplot2

&nbsp;&nbsp;&nbsp;**Data munging/wrangling -** dplyr

&nbsp;&nbsp;&nbsp;**Reshaping (speading and gathering) data -** tidyr, tibble 

&nbsp;&nbsp;&nbsp;**Importing/exporting data -** readr

**2. Now can you name two functions that you've used form each package that you listed above for these tasks?**

&nbsp;&nbsp;&nbsp;**Plotting -**geom_boxplot(), geom_point().

&nbsp;&nbsp;&nbsp;**Data munging/wrangling** -  mutate(), select().
   
&nbsp;&nbsp;&nbsp;**Reshaping (speading and gathering) data -** spread(), gather().

&nbsp;&nbsp;&nbsp;**Importing/exporting data -** read_delim(), read_csv().

##<a name="sec3">3.</a> R Basics

**1. Fix this code with fewest number of changes possible to it works:**
```{r}
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
```

**2. Fix this code so it works:**
```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```
**3. Look at the code below and comment on what happened to the values in the vector:**
my_vector <- c(1, 2, '3', '4', 5)
my_vector

Answer: They all got converted to String because of c() method.

##<a name="sec4">4.</a> Data import/export

**1. Download the rail_trail.txt file from Canvas (in the Midterm Exam section) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result.**

```{r}
file_path <- '/Users/ehlansiddiqi/Downloads/rail_trail.txt'
rail_trail_data <- read_delim(file_path, delim = '|')
glimpse(rail_trail_data)
```

**2. Export the file into a comma-separated file and name it “rail_trail.csv”. Make sure you define the path correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another glimpse.**

```{r}
write_csv(rail_trail_data, '/Users/ehlansiddiqi/Downloads/rail_trail.csv')
rail_trail_csv <- read_csv('/Users/ehlansiddiqi/Downloads/rail_trail.csv')
glimpse(rail_trail_csv)
```

##<a name="sec5">5.</a> Visualization

**1. Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.**

* I would use MADAM PRESIDENT and bold both words.
* This is not tidy. Age range and gender are used in the same line. Although used different color, but it does not give a clear visual. The circular shapes are also not to proportion and they are overwhelming be bigger than the graph header (yes/no).
* Major flaw in this chart is that it does not specify how many Men and Women within the age range given has provided the answer. A simplified histogram with y axis being age range and Men Women in parrellel bar would make it more effective.

**2. Reproduce this graphic using the diamonds data set.**
```{r}
ggplot(data = diamonds) +
  geom_boxplot(aes(x = cut, y= carat, fill = color), position = 'identity')+
  xlab('CUT OF DIAMOND')+
  ylab('CARAT OF DIAMOND')+
  coord_flip()
```

**3. The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again.**

The graph was not showing carats for each color. It had an overlap. Replacing the identity will fix it.
```{r}
ggplot(data = diamonds) +
  geom_boxplot(aes(x = cut, y= carat, fill = color), position = 'dodge')+
  xlab('CUT OF DIAMOND')+
  ylab('CARAT OF DIAMOND')+ 
  coord_flip()
```

##<a name="sec6">6.</a> Data munging and wrangling

**1. Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy. Note: this data set is called table2 and is available in the tidyverse package. It should be ready for you to use after you’ve loaded the tidyverse package.**

This data is not tidy because the 'type' and 'count' column contains only key of the count value.
This can be made tidy using the spread() method
```{r}
spread(table2, type, count)
```

**2. Create a new column in the diamonds data set called price_per_carat that shows the price of each diamond per carat (hint: divide). Only show me the code, not the output.**
```{r}
diamonds$price_per_carat <- (diamonds$price / diamonds$carat)
```

**3. For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.**
```{r}
result <- diamonds %>% 
  filter(price > 10000, carat < 1.5)
result
```

  **Do the results make sense? Why?**
    
    No.The prices seems inconsistent with the qulity of the dimaonds. A lesser carat and cut seems to have higher price.
    
  **Do we need to be wary of any of these numbers? Why?**
    
    Higher carat and cut should lead to higer price. So these numbers may not be accurate.
    
##<a name="sec7">7.</a> EDA

**Take a look at the txhousing data set that is included with the ggplot2 package and answer these questions:**

**1. During what time period is this data from?**
```{r}
min(txhousing$year, na.rm=T)
max(txhousing$year, na.rm=T)
```
This is from the year 2000 to 2015

**2. How many cities are represented?**
```{r}
sum(!duplicated(txhousing$city))
```
46

**3. Which city, month and year had the highest number of sales?**
```{r}
answer <- txhousing %>% filter(sales == max(txhousing$sales, na.rm=T)) 
glimpse(answer$city)
```

**4. What kind of relationship do you think exists between the number of listings and the number of sales? Check your assumption and show your work.**

The relationship is 1 to 1.
```{r}
ggplot(data = txhousing, mapping = aes(x = listings, y = sales))+
  geom_boxplot()
```

**5. What proportion of sales is missing for each city?**


```{r}
sum(is.na(txhousing$sales))
sum(!is.na(txhousing$sales))
aggregate(txhousing$sales, by= list(city = txhousing$city), FUN=sum, na.rm = TRUE)
aggregate(txhousing$sales, by= list(city = txhousing$city), FUN=sum, na.rm = FALSE)

```

**6. Looking at only the cities and months with greater than 500 sales:**

  Are the distributions of the median sales price (column name median), when grouped by city, different? The same? Show your work.

They are not the same
```{r}
over500 <- txhousing %>% filter(sales > 500)
over500 %>% group_by(city) %>% summarize(mean = mean(sales))
over500 %>% group_by(median) %>% summarize(mean = mean(sales))
```

Any cities that stand out that you’d want to investigate further?

Cities with higher mean.

##<a name="sec8">8.</a> Git and Github
link provided above
