---
title: "COMPSCIX 415.2 Homework 3"
author: "Ehlan Siddiqi"
date: "6/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
```

## ** Exercise 3.8.1 **
## 1. What is the problem with this plot? how can you impove it
## ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
##    geom_point()

## Answer:
## There are overlapping points, therefore unable have proper visual representation. This can be improved by adding slight noise with jitter.
```{r 3.8.1-Q1}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point(position = "jitter")
```

## 2. What parameters to geom_jitter() control the amount of jittering?
## Answer:
## using position_jitter() we can pass width and hight
```{r 3.8.1-Q2}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point(position = position_jitter(width =  2, height = 3))
```

## 3. Compare and contrast geom_jitter() with geom_count()
## Answer: geom_count is usufull when we have discrete data and overplotting as it counts the number of observations at each location. geom_jitter adds noise to the overlapping data and combines smaller datasets. 
##
```{r 3.8.1-Q3 jitter}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_jitter()
```
```{r 3.8.1-Q3 count}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_count()
```

## 4. What's the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that deomnstrates it.
## Answer:
## Default poisition for geom_boxplot() is "dodge"
```{r 3.8.1-Q4 default}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot()
```
```{r 3.8.1-Q4 position identity}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot(position =  "identity")
```

## ** Exercise 3.9.1 **

## 2. What does lab() do? Read the documentation.
## Answer:
## labs() gives variuos options to modify lebels in diffrent axis and title. It takes parameters for lable, title and subtitle
```{r 3.9.1-Q2 labs}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot() +
  labs(y = "Highway", x = "Car Class", title ="Hightway mgp by class")
```


## 4. What does the plot below tell you about the releationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?
## ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
##  geom_point() +
##  geom_abline() +
##  coord_fixed()

## Answer:
## The geom_abline adds reference line to a plot for visual comparison. the coord_fixed , keeps a fixed angle to compare city mileage against highway.
```{r 3.9.1-Q4 geom_abline and coord_fixed}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_abline() +
  coord_fixed()
```

## ** Exercise 4.4.1 **

## 1. Why does this code not work
##    my_variable <- 10
##    my_varıable

## Answer:
## The varables are not the same, the "i" character is differnet, therefore the variable my_varıable was never initialized.

## 2. Tweek each of the follwing R commands so that they run correctly:
## Answer:
## By sequential complie erro follwing were fixed
## * "fliter" mothod corrected to "filter"
## * "="" assingment corrected to "==" comparison
## * "diamond"" variable changed to corrected "diamonds"
```{r 4.4.1-Q2 tweek code}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```

## ** Exercise 5.2.4 **

## 1 Find all flights that

```{r 5.2.4-Q1}
glimpse(flights)
```
## Answer:
##    Had an arrival delay of two or more hours
```{r 5.2.4-Q1 1}
filter(flights, arr_delay > 120)
```
##    Flew to Houston (IAH or HOU)
```{r 5.2.4-Q1 2}
filter(flights, dest %in% c("IAH", "HOU"))
```
##    Were operated by United, American, or Delta
```{r 5.2.4-Q1 3}
filter(flights, carrier %in% c("UA","AA", "DL"))
```
##    Departed in summer (July, August, and September)
```{r 5.2.4-Q1 4}
filter(flights, month >=7, month <=9)
```
##    Arrived more than two hours late, but didn’t leave late
```{r 5.2.4-Q1 5}
filter(flights, dep_delay < 1, arr_delay > 120)
```
##    Were delayed by at least an hour, but made up over 30 minutes in flight
```{r 5.2.4-Q1 6}
filter(flights, dep_delay >= 60, arr_delay <= 30)
```
##    Departed between midnight and 6am (inclusive)
```{r 5.2.4-Q1 7}
filter(flights, dep_time <=600 | dep_time >= 2400)
```

## 3 How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
## Answer:
## arr_time is also missing, which means these flight never took off and was probably cancelled.
```{r 5.2.4-Q3}
filter(flights, is.na(dep_time))
```

## 4 
## Why is N^0 not missing? 
## Answer: all value of x^0 is defaulted to 1
```{r}
NA ^ 0
```

## Why is NA | TRUE not missing?
## Answer: for an or condition of one condition is true it will return true.
```{r}
NA | TRUE
```
## Why is FALSE & NA not missing
## Answer: For any & condition if one condition is false the return will be false
```{r}
FALSE & NA
```

## Can you figure out the general rule? (NA*0 is a tricky conterexample!)
## Answer:

## ** Exercise 5.4.1 **
## 1 Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights
## Answer:
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, starts_with("dep_"), starts_with("arr_"))
select(flights, ends_with("_time"), ends_with("_delay"))
select(flights, contains("dep_"), contains("arr_"))
select(flights, matches("^(dep|arr)_(time|delay)$"))
```


## 3 What does the one_of() function do? WHy might it be helpful in conjunction with this vector
## Answer:
## one_of() is used with column names are vector or String, 
## this will throw an error:
## select(flights, vars)
## one_of() is used to remedy this.
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

