---
title: "COMPSCIX 415.2 Homework 4"
author: "Ehlan Siddiqi"
date: "7/03/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
```

## **Exercise 5.6.7**

# 2. Come up with another approach that will give you the same output as 
# not_cancelled %>% count (dest) and not_cancelled %>% count(tailnum, wt
# disantce) (without using count)

#Answer:
```{r 5.6.7-Q2}
not_cancelled <-
  flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))
# count(dest)
not_cancelled %>%
  group_by(dest) %>%
  summarise(n = n())

# count(tailnum, wt = distance)
not_cancelled %>%
  group_by(tailnum) %>%
  summarize(n= sum(distance))
```

# 4. Look at the number of cancelled flights per day. is there a pattern? Is the proportion of calcelled flights related to average delay?

# Answer
```{r 5.6.7.Q4}
flights %>% 
  group_by(day) %>%
  summarise(cancelled_flights = mean(is.na(dep_delay)),
            dep_delay_mn = mean(dep_delay, na.rm = T),
            arr_delay_mn = mean(arr_delay, na.rm = T)) %>%
  ggplot(aes(y = cancelled_flights)) +
  geom_point(aes(x = dep_delay_mn)) +
  geom_point(aes(x = arr_delay_mn)) +
  labs(x = "Delay Average", y = "Cancelled Flights")
```
# delays are proportionate to cancellation

# 6. What does the sort argument to count() do. WHen might you use it?
# Answer:
# sort will sort results of the count.
```{r 5.6.7.Q6}
flights %>%
  count(flight, sort = T)
```

## **Exercise 10.5**

# 1. How can you tell if an object is a tibble?
# Answer:
# If you are able to explicitly print, it's a tibble.
```{r 10.5-1}
print(mtcars)
```

# 2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?
# Answer
```{r 10.5-2}
#df <- data.frame(abc = 1, xyz = "a")

# df$x
# tibble will throw a warning, as partial matching is not allowed

#df[, "xyz"]
# no difference

#df[, c("abc", "xyc")]
# no difference
```
# Data frame may throw in consistent data will tibble will not.

# 3. If you have the name of a variable stored in an ojbect, e.g var <- "mpg", how can you extract the reference variable form a tibble?

```{r 10.5-3}
var <- "mpg"
as_tibble(mtcars)[var]
```

# 6. What option controls how many additional column names are printed at at the footer of a tibble?
# Answer
```{r 10.5-6}
m <- 10
n <- 20
options(tibble.print_max = n, tibble.print_min = m)
# if more than n rows, print m rows.
```

## **Exercise 12.3.3 **

# 2. Why does this code fail
# Answer
# cannot have number as a column name
```{r 12.3.3-Q2}
table4a %>%
  gather('1999', '2000', key = "year", value = "cases")
```

# 3. Why does spreading this tibble fail? How could you ada a new column to fix the problem?

# Answer: name and key are unique, cannot have two age option for same name.
```{r 12.3.3-Q3}
people <- tribble(
   ~name,              ~key,       ~value, 
  #------------------/-----------/-------
  "Phillips Woods",   "age",      45,
  "Phillips Woods",   "height",   186,
  "Phillips Woods",   "age",      50,
  "Jessica Cordero",  "age",      37,
  "Jessica Cordero",  "height",   156
)
people %>%
  mutate(unique = c(1, 2, 2, 3, 3)) %>%
  select(unique, everything()) %>%
  spread(key, value)
```

# 4.Tidy the simple tibble below. Do ou need to spread or gather it? What are the variables?
```{r 12.3.3-Q4}
preg <- tribble (
  ~pregnant, ~male, ~female,
  "yes",      NA,    10,
  "no",       20,   12
)
preg %>%
  gather(gender, values, -pregnant)
```

## **Exercise 12.4.3
# 1. What do the extra and fill argument do in separate()? Experiment with the various options for the following two toy datasets.
```{r 12.4.3-Q1}
tibble(x = c("a,b,c", "d,e,f", "h,i,j")) %>%
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>%
  separate(x, c("one", "two", "three"))

## fill adds option to deal with missing columns
tibble(x = c("a,b,c", "d,e,f", "h,i,j")) %>%
  separate(x, c("one", "two", "three", "four", fill = "warn"))
tibble(x = c("a,b,c", "d,e,f", "h,i,j")) %>%
  separate(x, c("one", "two", "three", "four", fill = "right"))
tibble(x = c("a,b,c", "d,e,f", "h,i,j")) %>%
  separate(x, c("one", "two", "three", "four", fill = "left"))

## extra drops or merges the extra columnt
tibble(x = c("a,b,c", "d,e,f", "h,i,j")) %>%
  separate(x, c("one", "two", "three", "four", extra = "warn"))
tibble(x = c("a,b,c", "d,e,f", "h,i,j")) %>%
  separate(x, c("one", "two", "three", "four", extra = "drop"))
tibble(x = c("a,b,c", "d,e,f", "h,i,j")) %>%
  separate(x, c("one", "two", "three", "four", extra = "merge"))

```

# 2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?
```{r 12.4.3-Q2}
## unite and separate creates new column and keeps the existing ones. remove actually removes the original column that was used for unite or separate
```

