---
title: "COMPSCIX 415.2 Homework 2"
author: "Ehlan Siddiqi"
date: "6/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
```

## **Excercise 3.2.4 **

## 1. Run ggplot(data = mpg). What do you see?
```{r 3.2.4-q1}
ggplot(data = mpg)
```

## Answer: I see a grey plot with no dots or data on it.

## 2. How many rows are in mpg? How many columns?
```{r 3.2.4-Q2}
dim(mpg)
```
## Answer: 234 Rows, 11 columns

## 3. What does the drv variable describe? Read the help for ?mpg to find out.
```{r 3.2.4-Q3}
?mpg
```
## Answer: f = front-wheel drive, r = rear wheel drive, 4 = 4wd

## 4. Make a scatterplot of hwy vs cyl
```{r 3.2.4-Q4}
ggplot(data = mpg) + geom_point (mapping = aes(x = hwy, y = cyl))
```
## 5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r 3.2.4-Q5}
ggplot(data = mpg) + geom_point (mapping = aes(x = class, y = drv))
```
## Answer: This graph is not usefull because class and drv attributes are categorical

## ** Excercise 3.3.1 **

## 1. What’s gone wrong with this code? Why are the points not blue?
## Answer: Color cannot be part of Aesthetic, correct code:
```{r 3.3.1-Q1}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```
## 2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
```{r r 3.3.1-Q2}
head(mpg)
```
## Answer: the head command will dispay the the continues vs categorical under each column
## categorical: manufacturer, model, trans, drv, fl, class

## 3.Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r 3.3.1-Q3}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = class, color = trans, size = hwy, shape = drv))
```
## Answser, different scale is shown.

## 4. What happens if you map the same variable to multiple aesthetics?
```{r 3.3.1-Q4}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cty))
```
## Answer: chart still displays both.

## 5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
`
```{r 3.3.1-Q5}
?geom_point
```
## Answer: stroke asthetics can modify the width of the border or any object.

## 6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?
```{r 3.3.1-Q6}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = class, color = displ < 5))
```
## Answer: plot still loads the aestehtic.

## ** Excercise 3.5.1 **

## 4. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

```{r 3.5.1-Q4}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

## Answer: Facet wraps groups up individiual dataset separately. With coloring,they are grouped togehter to have a side by side comparison. For larger data coloring would be a better option to group all data together for comparison, as there maybe too manuy boxes to compare and will become difficult.

## 5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol argument?

```{r 3.5.1-Q5}
?facet_wrap
```
## Answer: nrow, ncol are number for rows and column.  Layout of columns can also be controled by scales and shrink.
## grid of fixed, therefor number of row and column cannot be modified.

## ** Excercise 3.6.1 **

## 1 What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
## Answer: will use a geom_line 

## 2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
```{r 3.6.1-Q2}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

### 3. What does show.legend = FALSE do? What happens if you remove it?
```{r 3.6.1-Q3}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(show.legend = TRUE)
```
## Answer: no changes for FALSE as legend is not included by default

## 4. What does the se argument in geom_smooth() do?
## Answer: se - display confidence interval around smooth? (TRUE by default)

## 5 Will these two graphs look different? Why/why not?
```{r 3.6.1-Q5}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```
## Answer: they do not look different. ggplot being parent object, the child ojbect will inherit from parent.

## ** Excercise 3.7.1 **

## 1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

## Answer the default geom associated with stat_summary is pointrange.
```{r 3.7.1-Q1}
ggplot(data = diamonds) + 
  geom_pointrange(
    mapping = aes(x = cut, y = depth), 
    fun.ymin = min, 
    fun.ymax = max, 
    fun.y = median, 
    stat = "summary")
```



